---
title: "Dice 10,000"
author: "Peter Crawford"
date: "9/1/2020"
output: pdf_document
---

### Introduction:

Rules for the game Dice (10,000): https://en.wikipedia.org/wiki/Dice_10000#:~:text=The%20first%20player%20to%20score,exactly%2010%2C000%20without%20going%20over.

We play where you need to score 1,000 to get on the board and there is no point requirement for turns after getting on the board. We also play where you can't roll from the previous player's dice - you start with six fresh dice every time. The rest of the rules on the above wikipedia link seem to be how we usually play.

The main question to be answered here is what score should I aim for each turn that will increase my chances of reaching 10,000 fastest? (i.e. should I always stop after rolling past 400? 500? 600?)


### Code:

Packages:
```{r}
library(tidyverse)
```

Create a function to roll dice. (Only variable being the number of dice rolled)
```{r}
roll <- function(num_of_dice) {
  ## RES: just code comment, 1:6 instead of c(1:6)
  sample(1:6, num_of_dice, replace = TRUE)
}
```

Create a function for scoring the rolled dice, and re-rolling if threshold not breached.
```{r}
## RES: code as it stands function doesn't run, is this function supposed to be "score2"?
score <- function(hand, threshold_value, start_points) {
  
  ## Starting values:
  values <- as.data.frame(table(hand)) ## get frequency
  values$hand <- as.numeric(levels(values$hand))[values$hand] 
  
  ## Straight:
  if (length(values$hand[values$Freq == 1]) == 6) { points <- 1500 }
  
  ## Combinations, ones, and fives:
  ## RES: since speed is an issue, for integers, isn't >= 3 the same as > 2
  values$combos <- ifelse(values$Freq > 2, (values$hand * 100) * (values$Freq - 2), 0)
  values$combos <- ifelse(values$combos == 100, values$combos * 10, values$combos)
  values$ones <- ifelse((values$combos == 0) & (values$hand == 1), 100 * values$Freq, 0)
  values$fives <- ifelse((values$combos == 0) & (values$hand == 5), 50 * values$Freq, 0)
  values$unused_dice <- ifelse((values$combos == 0) & (values$ones == 0) & (values$fives == 0), values$Freq, 0)
  
  ## Get the roll score, number of leftover dice, and running total score:
  points <- sum(values$combos) + sum(values$ones) + sum(values$fives)
  unused_dice <- sum(values$unused_dice)
  total_points <- start_points + points
  
  ## RES: is this even possible?
  if (points == 0) {
    return(points)
  }
  
  ## Determine whether or not to continue rolling:
  ## 1. Stop rolling if threshold is reached
  ## 2. Stop rolling if no points scored & no 
  
  ## RES: could remove a condition -- might be misunderstanding, but seems like the only time to roll more is if <threshold_value
  if ((total_points < threshold_value) | (unused_dice == 0)) {
    if (unused_dice == 0) {
      score2(hand = sample(1:6, 6, replace = TRUE), threshold_value = threshold_value, start_points = total_points)
    }
    else {
      score2(hand = sample(1:6, unused_dice, replace = TRUE), threshold_value = threshold_value, start_points = total_points)
    }
  }
  else {
    return(total_points)
  }
}
```

This function uses Monte Carlo simulation (replicate) to see how many turns taken to win with given threshold: 
```{r}
turns_to_win <- function(thresh_val) {
  
  total_score <- 0
  turn_scores <- NULL
  i <- 0
  
  while(total_score < 10000) {
    i <- i + 1
    ## RES: roll is kinda a useless function, due to memory shuffling, you would be better off moving it here
    current_roll <- sample(1:6, 6, replace = TRUE)
    turn_scores[i] <- score2(hand = current_roll, threshold_value = thresh_val, start_points = 0)
    total_score <- sum(turn_scores)
  }
  
  return(length(turn_scores))
}
```

Create a df with the threshold and the turns to win:
```{r}
n_reps <- 10 # When this value is greater than 10, computing time takes super long,
             # but the greater the reps, the closer we get to approximating the true values for each thresh.
             # Ideally, n_reps would be around 10,000 or so.

df_10000 <- data.frame("thresh" = 50, "ave_turns_to_win" = 20)

## RES: this is going to be slow because of the recursive nature of "score"
## try to parallelize around the inner loop where you are currently using replicate
create cluster
library(parallel)
cl <- makeCluster(detectCores()-1) ## on some computers this doesn't work as intended, just set it to number of cores-1 available
clusterExport(cl,c("score"))
for (i in 1:100) {
  
  temp_thresh <- 10 * i
  ## RES: Set a different seed on each member of the cluster (just in case)
  clusterSetRNGStream(cl)
  ## RES: need to play with this a bit to make sure it works, might need to export temp_thresh
  parSapply(cl, 1:10000, function(i,...) { turns_to_win(thresh_val = temp_thresh) } )
  turns_ave <- mean(replicate(n_reps, turns_to_win(thresh_val = temp_thresh)))  
  
  new_row <- c(temp_thresh, turns_ave)
  
 
  df_10000 <- rbind(df_10000, new_row)
  # print(i) # This is useful when n_reps is big as it will show us which iteration of our for loop we are on.
}

df_10000 <- df_10000[-1,] # Drop the initial value to 

## RES: stop the cluster
stopCluster(cl)
```

Plot, bottom of curve signifies faster at reaching 10,000 points (winning):
```{r}
ggplot(data = df_10000, aes(x = thresh, y = ave_turns_to_win)) +
  geom_line() + 
  geom_point() +
  xlab("Stopping Score (Threshold) for Each Turn") +
  ylab("Average Turns to Reach 10,000") +
  ggtitle("What target score should you aim for when playing Dice?") +
  scale_x_continuous(breaks = seq(0,1000, by = 100)) +
  # geom_smooth(method = "lm", se = FALSE)
  stat_smooth(method = "lm", se = TRUE, fill = NA, formula = y ~ poly(x, 2, raw = TRUE), colour = "blue")
```

